---
title: "A quick look at the first 2016 presidential debate"
author: "Megan Risdal"
date: "27 September 2016"
output:
  html_document:
    fig_width: 10
    fig_height: 6
    theme: spacelab
    highlight: kate
---

Just kidding with the Kernel title. Welcome to this brief analysis of the transcript from the first 2016 presidential debate!

```{r, message = F, warning = F}
# Load libraries
library(dplyr)
library(ggplot2)
library(gridExtra)
library(tidytext)

debate <- read.csv("../input/debate.csv", h =T, stringsAsFactors = F)
```

How were the speakers synced with respect to response length? Here I use word count against `Line` as a proxy for time.


```{r, message = F, warning = F}

debate <- debate %>% 
	mutate(word_count = sapply(gregexpr("[[:alpha:]]+", Text), function(x) sum(x > 0)))

ggplot(debate %>% filter(Speaker %in% c("Clinton", "Trump")), 
	aes(x = Line, y = word_count, fill = Speaker)) + 

	geom_bar(stat="identity") + 

	scale_x_continuous(breaks = seq(0, nrow(debate), 25)) +

	labs(x = "Time (Line in Transcript)", 
		 y = "Word Count in Candidate Response", 
		 title = "Candidates' Response Lengths over Time") +

	scale_fill_manual(values = c("#4169E1", "#E91D0E"))
```

What about simply looking at the distribution of their response lengths?

```{r, message = F, warning = F}

ggplot(debate %>% filter(Speaker %in% c("Clinton", "Trump")), 
	aes(x = Speaker, y = word_count, fill = Speaker)) +

	geom_boxplot() +

	#scale_y_log10() +

	labs(x = "Candidate", 
		 y = "Word Count in Candidate Responses", 
		 title = "Distribution of Candidates' Response Lengths") +

	scale_fill_manual(values = c("#4169E1", "#E91D0E"))

```

One thing I've wanted to try is tf-idf using tidy data principles as nicely demonstrated in Julia Silge's blog here: http://juliasilge.com/blog/Term-Frequency-tf-idf/ using her `tidytext` library.


```{r, message = F, warning = F}
# Calculate word frequencies
debate_words <- debate %>%
	filter(Speaker %in% c("Clinton", "Trump")) %>%
    unnest_tokens(word, Text) %>%
    count(Speaker, word, sort = TRUE) %>%
    ungroup()

# Calculate tf-idf
debate_words <- debate_words %>%
    bind_tf_idf(word, Speaker, n) %>%
    arrange(desc(tf_idf)) %>%
    mutate(word = factor(word, levels = rev(unique(word))))

# Create plots of the tf-idf for each candidate
# Clinton
clinton <- ggplot(debate_words %>% 
	   filter(Speaker == "Clinton") %>% 
	   top_n(10),
	aes(x = word, y = tf_idf)) +
	geom_bar(aes(alpha = tf_idf), 
		stat="identity", 
		fill = "#4169E1") +
	coord_flip() +
	#scale_y_continuous(limits = c(0, 0.002)) +
	labs(x = NULL, y = "tf-idf", title = "Clinton") +
	scale_alpha_continuous(range = c(0.6, 1), guide = FALSE)

# Trump
trump <- ggplot(debate_words %>% 
	   filter(Speaker == "Trump") %>% 
	   top_n(10),
	aes(x = word, y = tf_idf)) +	
	geom_bar(aes(alpha = tf_idf), 
		stat="identity", 
		fill = "#E91D0E") +
	coord_flip() +
	labs(x = NULL, y = "tf-idf", title = "Trump") +
	scale_alpha_continuous(range = c(0.6, 1), guide = FALSE)

# Plot
grid.arrange(clinton, trump, ncol = 2)
```

And there you have it! Hope you enjoy this dataset. Stay tuned for the next debate.